Decade of Lost Precision: A General Framework For Image Feature Matching Without Geometric Constraints

### 2014-11-28 paper.tex
Stefan's changes

In the literature two approaches to feature point matching have been pursued and later merged. The geometric approach tries to find unique keypoints in an image pair and matches them based on their location.  The descriptor centric approach on the other hand uses the local image information around the keypoint to create a descriptor and matches descriptors based on their similarity.

In the purely geometric approach, feature points are matched based on their position in the images. Scott and Longuet-Higgins \cite{scott1991algorithm} and Shapiro and Brady \cite{shapiro1992feature} introduced the use of spectral methods by deriving a coherent set of matches from the eigenvalues of the correspondence matrix. Other examples of this approach include \cite{sclaroff1995modal,carcassoni2003spectral}.

@@

A straightforward way to find a set of correspondences using only feature points is to apply a threshold to the similarity measure of the feature vectors, accepting only correspondences that score above a certain level of similarity \cite{szeliski2010}. When we match images with the assumption that the correspondence between two feature points will be unique, we can further increase precision by only matching a feature point to its nearest neighbor in terms of descriptor similarity. Instead of thresholding based on similarity, Deriche et al.~\cite{deriche1994robust} and Baumberg~\cite{baumberg2000reliable} propose using the ratio of the similarity of the best to second best correspondence of a given point to evaluate how unique it is. Their finding has later been tested by several independent teams, all concluding that thresholding based on this ratio is generally superior to thresholding based on similarity \cite{lowe2004sift,mikolajczyk2005performance,moreels2007evaluation,rabin2009statistical}.

@@

A plethora of solutions have combined ratio match with various geometric constraints to improve matching. These constraints are based on assumptions regarding the transformation between the query and \emph{target images}. At the stricter end we have epipolar constraints assuming that the images can be tied by a homography \cite{torr2000mlesac, chum2005matching} and angular constraints assuming the correspondences are angled similarly \cite{kim2008efficient, schmid1997local}. Often these approaches are made computationally feasible by modeling the case of feature correspondences as an instance of graph matching, where each feature is a vertex, and edge values correspond to a geometric relation between two features. In this case approximate graph matching algorithms can be used to efficiently establish an isomorphism between the graph of features in two images \cite{leordeanu2005spectral, torresani2008feature, yarkony2010covering, yuan2012efficient}. Others define image regions and reject or accept correspondences based on the regions they connect \cite{cho2009feature, wu2011robust}.

@@

In our previous paper \cite{arnfred2013mirror}, we introduced \emph{Mirror-Match} and \emph{Mirror-Match with Clustering}, two algorithms that we showed to perform better than the state of the art. The novel contributions of the present paper consists of presenting these together with several related algorithms in a general and comprehensive framework.  We further develop the theoretical foundations for comparing the algorithms and use these to formally prove a ranking in terms of performance for the different algorithms, which also enables us to understand why \emph{Mirror-Match} performs better than \emph{Ratio Match} in the first place.  In addition we benchmark all algorithms within the framework extensively on a much larger dataset consisting of over 3000 image pairs.

@@

\section{Matching Framework}
\label{S:MatchingMethods}
%
\subsection{Definitions}
%
The proposed framework is inspired by \emph{Ratio-Match} as introduced by Deriche et al.\ \cite{deriche1994robust} and later used by Baumberg~\cite{baumberg2000reliable} and Lowe~\cite{lowe2004sift}. \emph{Ratio-Match} is motivated by the observation that nearest neighbor matching is not necessarily the best strategy as illustrated by~\cite{lowe2004sift, mikolajczyk2005performance}. The distance between the feature descriptors of two nearest neighbors might tell us on a global level how much they resemble each other, but it does not tell us if other feature points are equally similar. \emph{Ratio-Match} makes use of the ratio between the nearest and second nearest neighbor as a heuristic for our confidence in the match. Matches are returned only if this ratio is lower than a given threshold $\tau$, filtering out nearest neighbors that are ambiguous because other points match almost equally well.

The underlying assumption in ratio match is that the point we seek to match in the \emph{query image} has only one true correspondence in a given \emph{target image} or no matches at all. In both cases we can infer that the \emph{second} nearest neighbor in the \emph{target image} is not a true correspondence. We look at the distance between the second nearest neighbor and the feature point as a \emph{baseline}. It tells us how similar the descriptors of two feature points can be when they are not a true correspondence. Some feature points might have very unique descriptors with large distances to false correspondences while others may be generic with plenty of similar points. Knowing the baseline for all features allows us to be lenient in the first case and cautious in the second. In practice \emph{Ratio-Match} scores a match by dividing the distance to the nearest neighbor with the distance to the second nearest neighbor (the baseline) to get an estimated measure for how distinct the correspondence is from a false match.

To make further discussion easier we introduce the following nomenclature:
\begin{itemize}
\item{Let $f_q$ be a feature point in the \emph{query image}}.
\item{Let $F$ be a set of features. $F_{t}$ denotes all
        features from the \emph{target image}}.
\item{Let $\tau \in [0 \ldots 1]$ be a threshold used to decide whether 
    to keep a match.} %We keep a match only if it has a \emph{uniqueness ratio} lower than $\tau$}
\item{Let the \emph{proposed match} be the nearest neighbor of a query 
    feature $f_q$ picked from a set of feature points that we will call 
the \emph{proposal set} which does not contain the query feature}.
\item{Let the \emph{baseline match} be the nearest neighbor of the query 
    feature $f_q$ picked from a set of feature points that we will call 
the \emph{baseline set} which neither contains the \emph{proposed match} 
nor the query feature}. In \emph{Ratio-Match} the \emph{baseline match} is the second nearest neighbor.
\end{itemize}

@@

\begin{figure}[t]
\centering
\includegraphics[width=0.95\columnwidth]{images/overview-wide}
\caption{Graphical representation of the \emph{Baseline} and 
\emph{Proposal Set} for different methods in the proposed framework}
\label{fig:overview}
\end{figure}

We can generalize \emph{Ratio-Match} by expanding on the idea of a \emph{baseline} and \emph{proposal set}. With \emph{Ratio-Match} these two sets are created from features in the \emph{target image}, but they need not be. If we use the features in the \emph{query image} as well as the combined features of both images we end up with six possible permutations of a \emph{Ratio-Match}-like algorithm. We illustrate these variations in Figure~\ref{fig:overview} and will go on to both prove theoretically and demonstrate empirically that \emph{Ratio-Match} is among the least performant of the pack.

The \emph{Mirror-Match} variation was originally introduced by the authors in~\cite{arnfred2013mirror} where it was shown to be superior to \emph{Ratio-Match}. In this paper we improve on this work by established formal proofs for the precision of the matching algorithms, by evaluating the algorithms on a more thorough dataset and by identifying \emph{Mirror-Match} as part of a general framework. We will not discuss the \emph{Self-Match-Ext} and \emph{Both-Match} variations in this paper since we prove them to be equivalent with \emph{Self-Match} and \emph{Mirror-Match} respectively in \ref{A:self} and \ref{A:mirror}.

@@

Algorithm~\ref{alg-gen} shows the generalized matching framework where the \emph{proposal set} and \emph{baseline set} are defined according to Figure~\ref{fig:overview}. Figure~\ref{fig:matching} illustrates the structure of this algorithm.

@@

For some methods it is possible that the \emph{proposal match} is a feature from the \emph{query image}, in which case we discard the match as an untrue correspondence.  It also happens that we encounter correspondences with $r>1$, in which case the match is also discarded.  Take for example the case of \emph{Self-Match} where we might find that the nearest neighbor of a feature in the \emph{target image} is further from the query feature than the nearest neighbor in the \emph{query image}. In this case the match is discarded.

@@

The main distinguishing factor between the algorithms in the framework is the final ratio between the distance of the two nearest neighbors of a query feature which decides if we keep or discard a match. We illustrate this process in Figure~\ref{fig:matching}.

We can define a calculation of this ratio that is common to all algorithms in the framework, which allows us to compare the algorithms theoretically. To do so we we introduce the \emph{uniqueness ratio} $r$, based on the concept of a nearest neighbor.

Given a feature $f_i$ and a set of features $F$, the nearest neighbor of $f_i$ in $F$ is calculated as follows: 
\begin{equation*}
\argmin_{f_j \in F} d(f_i, f_j)
\end{equation*}
Here, $d(f_i, f_j)$ is the distance between feature descriptors.With SIFT and SURF this is the Euclidean distance of the feature vectors \cite{lowe2004sift,bay2006surf}, whereas BRIEF, BRISK, and FREAK use the Hamming distance \cite{leutenegger2011brisk,calonder2010brief,alahi2012freak}.

If we let $f_p \in F_{p}$ and $f_b \in F_{b}$ be the nearest neighbors of a feature $f_q$ in $F_{p}$ and $F_{b}$, the \emph{uniqueness ratio} is defined as follows:
\begin{align*}
    r &= \text{r}(f_{q}, F_{p}, F_{b}) \\
        &= \frac{d(f_{q}, f_{p})}{d(f_{q}, f_{b})}.
\end{align*}

The performance in terms of precision and recall of any algorithm from the proposed framework is uniquely identified by the \emph{uniqueness ratio} $r$. To show this we let $K$ be the number of possible true correspondences between \emph{query} and \emph{target image}. \emph{Precision} and \emph{recall} are defined as:
\begin{align*}
    \textrm{Recall} &= \frac{\#\textrm{Correct}}{K} \\
    \textrm{Precision} &= \frac{\#\textrm{Correct}}{\#\textrm{Correct} + \#\textrm{Incorrect}}
\end{align*}
Given $f_{p} \in F_{p}$ as the nearest neighbor of some $f_q \in F_{q}$, we define the set of features $F_{true}$ as $\left\{ f_{q} \in F_{q} \mid f_{p} \text{ is a true correspondence of } f_{q} \right\}$ and $F_{false}$ as $F_{q} \setminus F_{true}$.  We can then define the number of correct and incorrect matches as follows:
\begin{align*}
    \textrm{keep\_match}(f_{q}) &= \twopartdef{ 1 }{\text{r}(f_{q}, 
    F_{p}, F_{b}) <
    \tau}{0}{\textrm{otherwise}} \\
    \#\textrm{Correct} &= \sum_{f_{q} \in F_{true}} 
    \textrm{keep\_match}(f_{q})\\
    %\mathbbm{1}_{\left\{ \right\}} \\
    \#\textrm{Incorrect} &= \sum_{f_{q} \in F_{false}}
    \textrm{keep\_match}(f_{q})
    %\mathbbm{1}_{\left\{ r < \tau\right\}}
\end{align*}
If for every query feature $f_{q}$, $r$ is identical across two matching methods, then they return identical results.

@@

Under the following three assumptions, which largely reflect the performance of matching feature points in practice, we can theoretically compare the performance of the different algorithms shown in Figure~\ref{fig:overview}:

\begin{enumerate}
    \item{For a query and \emph{target image} we assume that for any point in 
        the \emph{query image} there is at most one real correspondence in the 
    \emph{target image} and no real correspondences in the \emph{query image} itself 
as assumed by \emph{Ratio-Match}.}
    \item{We assume that the distance between two feature descriptors within the \emph{query image} is larger than their distance to a real correspondence in the \emph{target image}.  More precisely, given $f_q$, a feature from the \emph{query image} for which a real correspondence, $f_{match}$, exists in the \emph{target image}, we assume that $\forall f_i \in F_{q}: d(f_q,f_{match}) < d(f_q, f_i)$.}
    %\item{We assume that features within one image resemble each other 
            %more than features in different images. That is, given 
            %$f_a,
    %        f_i \in F_1$ and $f_b, f_j \in F_2$ where $(f_i, f_j)$ and 
    %    $(f_a, f_b)$ are true correspondences, and $F_1$ and $F_2$ are 
    %features collected from different images then we expect that 
            %$d(f_a, f_i)% < d(f_a, f_j)$ and $d(f_b, f_j) < d(f_b, 
            %f_i)$}
     \item{For a set of query features $F_{q}$ with a true 
             correspondence in the \emph{target image}, we assume that 
             the distribution of \emph{uniqueness ratios} using $F_{t}$ as 
             the baseline set is similar to using $F_{q}$, since the 
         two images in the case of true correspondences are bound 
     to share part of the same scene.}
    \end{enumerate}

Based on these assumptions we prove that \emph{Ratio-Match-Ext} is equal
or better than \emph{Ratio-Match} in terms of both precision and recall.  Consider the nearest neighbor $f_{p}$ of a query feature $f_{q}$ and the two cases where we either have the $f_{p} \in F_{q}$ or $f_{p} \in F_{t}$. For $f_{p} \in F_{t}$ the uniqueness ratio of \emph{Ratio-Match-Ext} is:
\begin{align*}
    r &= \text{r}(f_{q}, F_{p}, F_{b}) \\
        &= \text{r}(f_{q}, F_{q} \cup F_{t}, F_{b})\\
        &= \text{r}(f_{q}, F_{t}, F_{b})
\end{align*}
Since $F_{p} = F_{t}$ for \emph{Ratio-Match} this shows that when the nearest neighbor is found in the \emph{target image}, the two algorithms behave identically. For $f_{p} \in F_{q}$ \emph{Ratio-Match} gives us the following ratio:
\begin{align*}
    r &= \text{r}(f_{q}, F_{p}, F_{b}) \\
        &= \text{r}(f_{q}, F_{t}, F_{t}).
\end{align*}

That is, \emph{Ratio-Match} calculates the ratio based on the two nearest correspondences in the \emph{target image}. \emph{Ratio-Match-Ext} on the other hand does not return any correspondence because the nearest neighbor is in the \emph{query image}. Since query features with a nearest neighbor in the \emph{query image} are false correspondences per assumption \#1 and \#2, this proves that \emph{Ratio-Match-Ext} has superior \emph{Precision} to \emph{Ratio-Match} while maintaining equal \emph{Recall}.

Next we show that \emph{Mirror-Match} is equal or better than \emph{Ratio-Match-Ext} in terms of precision and recall. Consider as before the nearest neighbor $f_{p}$ of a query feature $f_{q}$. When $f_{p}$ resides in the \emph{query image} both algorithms behave alike and discard the match.  However, consider the uniqueness ratio of \emph{Mirror-Match} for the case where $f_{p}$ resides in the \emph{target image}:
\begin{align*}
    r &= \text{r}(f_{q}, F_{p}, F_{b}) \\
        &= \text{r}(f_{q}, F_{p}, F_{q} \cup F_{t})\\
        &= \max( \text{r}(f_{q}, F_{p}, F_{t}), 
    \text{r}(f_{q}, F_{p}, F_{q}) ).
\end{align*}
For a true correspondence the \emph{uniqueness ratio} using $F_{q}$ as a baseline is distributed similarly to the ratio when using $F_{t}$ according to assumption \#3. In this case the algorithm performs like \emph{Ratio-Match-Ext}.  However, for an untrue correspondence with a baseline match in the \emph{query image} which is closer than the baseline match in the \emph{target image}, \emph{Mirror Match} will return a worse \emph{uniqueness ratio}. This in turns means that \emph{Mirror Match} is equal or better than \emph{Ratio-Match-Ext} in terms of  \emph{Precision} while maintaining equal \emph{Recall}.

Using a similar procedure we can prove that \emph{Self-Match} is equal to \emph{Self-Match-Ext} and \emph{Both-Match} is equal to \emph{Mirror Match}.  The details can be found in \ref{A:self} and \ref{A:mirror}, respectively.

@@

\emph{Self-Match} is most closely related to \emph{Ratio-Match} in that both the \emph{baseline} and \emph{proposal} set are created based on only one image. While the \emph{proposal set} for both algorithms is based on the \emph{target image}, \emph{Self-Match} uses the \emph{query image} for the \emph{baseline set} while \emph{Ratio-Match} sticks with the \emph{target image}. For cases where there is a lot of overlap in between the \emph{target} and \emph{query image} they should perform similarly. However when we match images where the \emph{target image} may not overlap at all with the \emph{query image}, using the \emph{query image} and not the \emph{target image} as a \emph{baseline set} seems like a reasonable choice given that the baseline set in this case would be closer to the feature matched and more strictly rule out untrue correspondences.

@@

The first assumption states that there is at most one unique match to every feature point, which is often the case for natural images. However, for the recognition of object classes or in images with repetitive content, a point in the \emph{query image} might have several true correspondences in the \emph{target image}, so using a \emph{baseline match} from the \emph{target image} will lead to a much higher \emph{uniqueness ratio}. However, as long as the \emph{query image} does not contain repetitive objects, we can still use \emph{Self-Match} without loss of precision. Conversely the query image could have several similar objects which could negatively impact performance for all algorithms but \emph{Ratio-Match} in cases where the target image only contained one such object. However when applied to scene matching we would expect there to be several similar objects in the target image too for cases of matching scenes. In this case it could easily be ambiguous to match any of these points together. Here the uniqueness assumption plays to our advantage by forcing us to ignore image features that will lead to ambiguous matches.

According to the second assumption, feature descriptors behave such that when matched, a true correspondence to a \emph{query feature} will always be closer in distance than any other feature in the query image. For this assumption to be violated either the \emph{query image} contains repeating patterns (violating assumption \#1) or the two images are different enough (i.e.\ taken from very different viewpoints or under different lighting conditions) that the descriptor distance between two unrelated features ends up being closer than the distance between a feature and its true correspondence. In this case \emph{Ratio-Match-Ext} and \emph{Mirror-Match} would discard the match while \emph{Ratio-Match} would still attempt to match the descriptor to a feature in the \emph{target image}.

The empirical results of \emph{Self-Match} in Figure~\ref{fig:all_objects} and Figure~\ref{fig:patches} support the assumption. \emph{Self-Match} discards any match where a feature in the query image is more similar to the query feature, yet the algorithm consistently performs equally well or better than \emph{Ratio-Match} supporting the notion that for almost all true correspondences the best match is found in the \emph{target image}. As we would expect this assumption breaks down as we increase $\tau$ to accept matches with higher uniqueness ratios, illustrated by the higher recall rates of \emph{Ratio-Match} compared to \emph{Self-Match} in most plots. Once we accept almost all proposed nearest neighbors as matches we are bound to encounter more non-distinct feature points for which even true correspondences might have better matches amongst the \emph{query features}.

>>>>

In the literature two approaches to feature point matching have been pursued and later merged, namely the \emph{geometric approach} and the \emph{descriptor-centric approach}.
%The geometric approach tries to find unique keypoints in an image pair and matches them based on their location.  The descriptor centric approach on the other hand uses the local image information around the keypoint to create a descriptor and matches descriptors based on their similarity.

In the purely geometric approach, feature points are matched based on their location in the images. Scott and Longuet-Higgins \cite{scott1991algorithm} and Shapiro and Brady \cite{shapiro1992feature} introduced the use of spectral methods by deriving a coherent set of matches from the eigenvalues of the correspondence matrix. Other examples of this approach include \cite{sclaroff1995modal,carcassoni2003spectral}.

@@

A straightforward way to find a set of correspondences using only feature points is to apply a threshold to the similarity measure of the feature vectors, accepting only correspondences that score above a certain level of similarity \cite{szeliski2010}. When we match images with the assumption that the correspondence between two feature points will be unique, we can further increase precision by only matching a feature point to its nearest neighbor in terms of descriptor similarity. Instead of thresholding based on similarity, Deriche et al.~\cite{deriche1994robust} and Baumberg~\cite{baumberg2000reliable} propose using the \emph{ratio} of the similarity of the best to second best correspondence of a given point to evaluate how unique it is. Their finding has later been tested by several independent teams, all concluding that thresholding based on this ratio is generally superior to thresholding based on similarity \cite{lowe2004sift,mikolajczyk2005performance,moreels2007evaluation,rabin2009statistical}.

@@

A plethora of solutions have combined ratio match with various geometric constraints to improve matching. These constraints are based on assumptions regarding the transformation between the query and \emph{target images}. At the stricter end we have epipolar constraints, assuming that the images can be tied by a homography \cite{torr2000mlesac, chum2005matching}, and angular constraints, assuming the correspondences are angled similarly \cite{kim2008efficient, schmid1997local}. Often these approaches are made computationally feasible by modeling feature correspondences as an instance of graph matching, where each feature is a vertex, and edge values correspond to a geometric relation between two features. Approximate graph matching algorithms can then be used to efficiently establish an isomorphism between the graph of features in two images \cite{leordeanu2005spectral, torresani2008feature, yarkony2010covering, yuan2012efficient}. Others define image regions and reject or accept correspondences based on the regions they connect \cite{cho2009feature, wu2011robust}.

@@

In our previous paper \cite{arnfred2013mirror}, we introduced \emph{Mirror-Match} and \emph{Mirror-Match with Clustering}, two algorithms that outperform the state of the art. The novel contributions of the present paper consist of presenting  these algorithms together with several related existing algorithms in a general and comprehensive framework.  We further develop the theoretical foundations for comparing the algorithms and use these to formally prove a ranking in terms of performance for the different algorithms. This also enables us to understand why \emph{Mirror-Match} performs better than \emph{Ratio Match} in the first place.  In addition we benchmark all algorithms within the framework extensively on a much larger dataset containing over 3000 image pairs.

@@

\section{Matching Framework}
\label{S:MatchingMethods}
%
\subsection{Definitions}
%
The proposed framework is inspired by \emph{Ratio-Match} as introduced by Deriche et al.\ \cite{deriche1994robust} and later used by Baumberg~\cite{baumberg2000reliable} and Lowe~\cite{lowe2004sift}. \emph{Ratio-Match} is motivated by the observation that nearest neighbor feature matching is not necessarily the best strategy as illustrated by~\cite{lowe2004sift, mikolajczyk2005performance}. The distance between the feature descriptors of two nearest neighbors might tell us on a global level how much they resemble each other, but it does not tell us if other feature points are equally similar. \emph{Ratio-Match} makes use of the ratio between the nearest and second nearest neighbor as a heuristic to determine the confidence of the match. Matches are returned only if this ratio is lower than a given threshold $\tau$, filtering out feature points that are ambiguous because others match almost equally well.

The underlying assumption in ratio match is that the point we seek to match in the \emph{query image} has only one true correspondence in a given \emph{target image} or no matches at all. In both cases we can infer that the \emph{second} nearest neighbor in the \emph{target image} is not a true correspondence. We look at the distance between the second nearest neighbor and the feature point as a \emph{baseline}. It tells us how similar the descriptors of two feature points can be when they are not a true correspondence. Some feature points might have very unique descriptors with large distances to false correspondences, while others may be generic with plenty of similar points. Knowing the baseline for all features allows us to be lenient in the first case and cautious in the second. In practice \emph{Ratio-Match} scores a match by dividing the distance to the nearest neighbor with the distance to the second nearest neighbor (the baseline) to estimate how distinct the correspondence is from a false match.

In what follows we will use the following nomenclature:
\begin{itemize}
\item{Let $f_q$ be a feature point in the \emph{query image}}.
\item{Let $F$ be a set of features. $F_{t}$ denotes all features from the \emph{target image}}.
\item{Let $\tau \in [0 \ldots 1]$ be a threshold used to decide whether to keep a match.} %We keep a match only if it has a \emph{uniqueness ratio} lower than $\tau$}
\item{Let the \emph{proposed match} be the nearest neighbor of a query 
    feature $f_q$ picked from a set of feature points that we will call 
the \emph{proposal set}, which does not contain the query feature}.
\item{Let the \emph{baseline match} be the nearest neighbor of the query 
    feature $f_q$ picked from a set of feature points that we will call 
the \emph{baseline set}, which contains neither the \emph{proposed match} 
nor the query feature}. In \emph{Ratio-Match} the \emph{baseline match} is the second nearest neighbor.
\end{itemize}

@@

\begin{figure}[t]
\centering
\includegraphics[width=0.95\columnwidth]{images/overview-wide}
\caption{Graphical representation of the \emph{baseline set} and \emph{proposal set} for different methods in the proposed framework.}
\label{fig:overview}
\end{figure}

We can generalize \emph{Ratio-Match} by expanding on the idea of a \emph{baseline} and \emph{proposal set}. With \emph{Ratio-Match} these two sets are created from features in the \emph{target image}, but they need not be. If we use the features in the \emph{query image} as well as the combined features of both images, we end up with six possible permutations of a \emph{Ratio-Match}-like algorithm. We illustrate these variants in Figure~\ref{fig:overview} and will go on to both prove theoretically and demonstrate empirically that \emph{Ratio-Match} is among the least performant of the pack.

The \emph{Mirror-Match} variant was originally introduced by the authors in~\cite{arnfred2013mirror}, where it was shown to be superior to \emph{Ratio-Match}. %In this paper we improve on this work by establishing formal proofs for the precision of the matching algorithms, by evaluating the algorithms on a more thorough dataset and by identifying \emph{Mirror-Match} as part of a general framework.

We will not discuss the \emph{Self-Match-Ext} and \emph{Both-Match} variants in this paper since we prove them to be equivalent with \emph{Self-Match} and \emph{Mirror-Match} respectively in \ref{A:self} and \ref{A:mirror}.

@@

Algorithm~\ref{alg-gen} shows the generalized matching framework, where the \emph{proposal set} and \emph{baseline set} are defined according to Figure~\ref{fig:overview}. Figure~\ref{fig:matching} illustrates the structure of this algorithm.

@@

For some methods it is possible that the \emph{proposal match} is a feature from the \emph{query image}, in which case we discard the match as a false correspondence.  It also happens that we encounter correspondences with $r>1$, in which case the match is also discarded.  Take for example the case of \emph{Self-Match} where we might find that the nearest neighbor of a feature in the \emph{target image} is further from the query feature than the nearest neighbor in the \emph{query image}. In this case the match is discarded.

@@

The main distinguishing factor between the algorithms in the framework is the final ratio between the distance of the two nearest neighbors of a query feature, which determines if we keep or discard a match. We illustrate this process in Figure~\ref{fig:matching}.

We can define a calculation of this ratio that is common to all algorithms in the framework, which allows us to compare the algorithms theoretically. To do so, we introduce the \emph{uniqueness ratio} $r$, based on the concept of a nearest neighbor.

Given a feature $f_i$ and a set of features $F$, the nearest neighbor of $f_i$ in $F$ is calculated as follows: 
\begin{equation*}
\argmin_{f_j \in F} d(f_i, f_j)
\end{equation*}
Here, $d(f_i, f_j)$ is the distance between feature descriptors. With SIFT and SURF this is the Euclidean distance of the feature vectors \cite{lowe2004sift,bay2006surf}, whereas BRIEF, BRISK, and FREAK use the Hamming distance \cite{leutenegger2011brisk,calonder2010brief,alahi2012freak}.

Let $f_p \in F_{p}$ and $f_b \in F_{b}$ be the nearest neighbors of a feature $f_q$ in $F_{p}$ and $F_{b}$.  The \emph{uniqueness ratio} is defined as follows:
\begin{align*}
    r &= \text{r}(f_{q}, F_{p}, F_{b}) \\
        &= \frac{d(f_{q}, f_{p})}{d(f_{q}, f_{b})}.
\end{align*}

The performance in terms of precision and recall of any algorithm in the proposed framework is uniquely identified by the \emph{uniqueness ratio} $r$. To show this, let $K$ be the number of possible true correspondences between \emph{query} and \emph{target image}. \emph{Precision} and \emph{recall} are defined as:
\begin{align*}
    \textrm{Recall} &= \frac{\#\textrm{Correct}}{K} \\
    \textrm{Precision} &= \frac{\#\textrm{Correct}}{\#\textrm{Correct} + \#\textrm{Incorrect}}
\end{align*}
Given $f_{p} \in F_{p}$ as the nearest neighbor of some $f_q \in F_{q}$, we define the set of features $F_{true}$ as $\left\{ f_{q} \in F_{q} \mid f_{p} \text{ is a true correspondence of } f_{q} \right\}$ and $F_{false}$ as $F_{q} \setminus F_{true}$.  We can then define the number of correct and incorrect matches as follows:
\begin{align*}
    \textrm{keep\_match}(f_{q}) &= \twopartdef{ 1 }{\text{r}(f_{q}, 
    F_{p}, F_{b}) <
    \tau}{0}{\textrm{otherwise}} \\
    \#\textrm{Correct} &= \sum_{f_{q} \in F_{true}} 
    \textrm{keep\_match}(f_{q})\\
    %\mathbbm{1}_{\left\{ \right\}} \\
    \#\textrm{Incorrect} &= \sum_{f_{q} \in F_{false}}
    \textrm{keep\_match}(f_{q})
    %\mathbbm{1}_{\left\{ r < \tau\right\}}
\end{align*}
If for every query feature $f_{q}$, $r$ is identical across two matching methods, then they return identical results.

@@

Under the following three assumptions, which largely reflect the performance of matching feature points in practice, we can theoretically compare the performance of the different algorithms shown in Figure~\ref{fig:overview}:
\begin{enumerate}
    \item{For any point in 
        the \emph{query image} there is at most one real correspondence in the 
    \emph{target image} and no real correspondences in the \emph{query image} itself, 
as assumed by \emph{Ratio-Match}.}
    \item{The distance between two feature descriptors within the \emph{query image} is larger than their distance to a real correspondence in the \emph{target image}.  More precisely, given $f_q$, a feature from the \emph{query image} for which a real correspondence, $f_{match}$, exists in the \emph{target image}, we assume that $\forall f_i \in F_{q}: d(f_q,f_{match}) < d(f_q, f_i)$.}
    %\item{We assume that features within one image resemble each other 
            %more than features in different images. That is, given 
            %$f_a,
    %        f_i \in F_1$ and $f_b, f_j \in F_2$ where $(f_i, f_j)$ and 
    %    $(f_a, f_b)$ are true correspondences, and $F_1$ and $F_2$ are 
    %features collected from different images then we expect that 
            %$d(f_a, f_i)% < d(f_a, f_j)$ and $d(f_b, f_j) < d(f_b, 
            %f_i)$}
     \item{For a set of query features $F_{q}$ with a true 
             correspondence in the \emph{target image}, 
             the distribution of \emph{uniqueness ratios} using $F_{t}$ as 
             the baseline set is similar to using $F_{q}$, since the 
         two images in the case of true correspondences are bound 
     to share part of the same scene.}
    \end{enumerate}

Based on these assumptions we prove that \emph{Ratio-Match-Ext} is equal to or better than \emph{Ratio-Match} in terms of both precision and recall.  Consider the nearest neighbor $f_{p}$ of a query feature $f_{q}$ and the two cases where we either have $f_{p} \in F_{q}$ or $f_{p} \in F_{t}$. For $f_{p} \in F_{t}$ the uniqueness ratio of \emph{Ratio-Match-Ext} is:
\begin{align*}
    r &= \text{r}(f_{q}, F_{p}, F_{b}) \\
        &= \text{r}(f_{q}, F_{q} \cup F_{t}, F_{b})\\
        &= \text{r}(f_{q}, F_{t}, F_{b})
\end{align*}
Since $F_{p} = F_{t}$ for \emph{Ratio-Match}, when the nearest neighbor is found in the \emph{target image}, the two algorithms behave identically. For $f_{p} \in F_{q}$ \emph{Ratio-Match} gives us the following ratio:
\begin{align*}
    r &= \text{r}(f_{q}, F_{p}, F_{b}) \\
        &= \text{r}(f_{q}, F_{t}, F_{t}).
\end{align*}

That is, \emph{Ratio-Match} calculates the ratio based on the two nearest correspondences in the \emph{target image}. \emph{Ratio-Match-Ext} on the other hand does not return any correspondence, because the nearest neighbor is in the \emph{query image}. Since query features with a nearest neighbor in the \emph{query image} are false correspondences per assumption \#1 and \#2, this proves that \emph{Ratio-Match-Ext} has superior \emph{Precision} to \emph{Ratio-Match} while maintaining equal \emph{Recall}.

Next we show that \emph{Mirror-Match} is equal or better than \emph{Ratio-Match-Ext} in terms of precision and recall. Consider as before the nearest neighbor $f_{p}$ of a query feature $f_{q}$. When $f_{p}$ resides in the \emph{query image}, both algorithms behave alike and discard the match.  However, consider the uniqueness ratio of \emph{Mirror-Match} for the case where $f_{p}$ resides in the \emph{target image}:
\begin{align*}
    r &= \text{r}(f_{q}, F_{p}, F_{b}) \\
        &= \text{r}(f_{q}, F_{p}, F_{q} \cup F_{t})\\
        &= \max( \text{r}(f_{q}, F_{p}, F_{t}), 
    \text{r}(f_{q}, F_{p}, F_{q}) ).
\end{align*}
For a true correspondence the \emph{uniqueness ratio} using $F_{q}$ as a baseline is distributed similarly to the ratio when using $F_{t}$ according to assumption \#3. In this case the algorithm performs like \emph{Ratio-Match-Ext}.  However, for an untrue correspondence with a baseline match in the \emph{query image} which is closer than the baseline match in the \emph{target image}, \emph{Mirror Match} will return a worse \emph{uniqueness ratio}. This in turns means that \emph{Mirror Match} is equal or better than \emph{Ratio-Match-Ext} in terms of  \emph{Precision} while maintaining equal \emph{Recall}.

Using a similar procedure we can prove that \emph{Self-Match} is equal to \emph{Self-Match-Ext}, and \emph{Both-Match} is equal to \emph{Mirror Match}.  The details can be found in \ref{A:self} and \ref{A:mirror}, respectively.

@@

\emph{Self-Match} is most closely related to \emph{Ratio-Match} in that both the \emph{baseline} and \emph{proposal} set are created based on only one image. While the \emph{proposal set} for both algorithms is based on the \emph{target image}, \emph{Self-Match} uses the \emph{query image} for the \emph{baseline set}, whereas \emph{Ratio-Match} sticks with the \emph{target image}. For cases with a lot of overlap between the \emph{target} and \emph{query image} they should perform similarly. However when we match images where the \emph{target image} may not overlap at all with the \emph{query image}, using the \emph{query image} and not the \emph{target image} as a \emph{baseline set} seems like a reasonable choice, given that the baseline set would be closer to the feature matched in this case and more strictly rule out false correspondences.

@@

The first assumption states that there is at most one unique match to every feature point, which is often the case for natural images. However, for the recognition of object classes or in images with repetitive content, a point in the \emph{query image} might have several true correspondences in the \emph{target image}, so using a \emph{baseline match} from the \emph{target image} will lead to a much higher \emph{uniqueness ratio}. However, as long as the \emph{query image} does not contain repetitive objects, we can still use \emph{Self-Match} without loss of precision. Conversely the query image could have several similar objects which could negatively impact performance for all algorithms except \emph{Ratio-Match} in cases where the target image only contains one such object. However when applied to scene matching we would normally expect there to be several similar objects in the target image too. In this case it could easily be ambiguous to match any of these points together. Here the uniqueness assumption plays to our advantage by forcing us to ignore features that will lead to ambiguous matches.

According to the second assumption, feature descriptors behave such that when matched, a true correspondence to a \emph{query feature} will always be closer in distance than any other feature in the query image. For this assumption to be violated, either the \emph{query image} has to contain repeating patterns (violating assumption \#1), or the two images are so different (i.e.\ taken from very different viewpoints or under different lighting conditions) that the descriptor distance between two unrelated features ends up being closer than the distance between a feature and its true correspondence. In this case \emph{Ratio-Match-Ext} and \emph{Mirror-Match} would discard the match while \emph{Ratio-Match} would still attempt to match the descriptor to a feature in the \emph{target image}.

The empirical results of \emph{Self-Match} in Figure~\ref{fig:all_objects} and Figure~\ref{fig:patches} support the second assumption. \emph{Self-Match} discards any match where a feature in the query image is more similar to the query feature, yet the algorithm consistently performs equally well or better than \emph{Ratio-Match}, providing further evidence to the notion that for almost all true correspondences the best match is found in the \emph{target image}. As we would expect this assumption breaks down as we increase $\tau$ to accept matches with higher uniqueness ratios, illustrated by the higher recall rates of \emph{Ratio-Match} compared to \emph{Self-Match} in most plots. Once we accept almost all proposed nearest neighbors as matches, we are bound to encounter more non-distinct feature points for which even true correspondences might have better matches amongst the \emph{query features}.
